### 2017-05-26 fieldJacobi.tex
Updating illustrations

\documentclass[border=10pt, tikz]{standalone}
\usepackage{tikz}
\usepackage{tikz-3dplot}
        \usetikzlibrary{patterns,hobby}

\tdplotsetmaincoords{110}{50}

\tikzset{%
  % Specifications for style of arrows:
   M/.style={
           -,
           red,},
   F/.style={
           -,
           blue,},
   W/.style={
           -,
           green,},
   O/.style={
           -,
           gray,},
   hidden/.style={
   			dashed,},                            
}

\def \zlevel {1.25}

\begin{document}

\begin{tikzpicture}[scale=2, tdplot_main_coords,thick]

\node (O) at (0,0,0)  {};
\node (A) at (0,3,0)  {};
\node (B) at (3,3,0)  {};
\node (C) at (3,0,0)  {};

\draw[M] (O.center) -- node [below] {$M$} (A.center);

\draw[F] (A.center) --  node [below right] {$F$} (B.center);
\draw[F,hidden] (B.center) --  node [right] {} (C.center); 
\draw[F,hidden] (C.center) --  node [right] {} (O.center);  
\path[fill=blue!5,opacity=.5] (O.center) to (A.center) to (B.center) to (C.center) to (O.center);

\node (O1) at (0,0,\zlevel)  {};
\node (A1) at (0,3,\zlevel)  {};
\node (B1) at (3,3,\zlevel)  {};
\node (C1) at (3,0,\zlevel)  {};

\draw[O] (O1.center) -- node [below] {$0(F)$} (A1.center); 
\draw[O] (A1.center) --  node [right] {} (B1.center);
\draw[O,hidden] (B1.center) --  node [right] {} (C1.center); 
\draw[O,hidden] (C1.center) --  node [right] {} (O1.center);

\node (O2) at (0,0,3)  {};
\node (A2) at (0,3,3)  {};
\node (B2) at (3,3,3)  {};
\node (C2) at (3,0,3)  {};

\draw[W] (O2.center) -- node [below] {} (A2.center); 
\draw[W] (A2.center) --  node [right] {} (B2.center);
\draw[W] (B2.center) --  node [right] {} (C2.center); 
\draw[W] (C2.center) --  node [above] {$W$} (O2.center);

\draw[W] (O.center) -- node [below] {} (O2.center); 
\draw[W] (A.center) -- node [below] {} (A2.center); 
\draw[W] (B.center) -- node [below] {} (B2.center); 
\draw[W,hidden] (C.center) -- node [below] {} (C2.center);

\path[fill=green!5,opacity=.25] (O.center) to (C.center) to (C2.center) to (O2.center) to (O.center);
\path[fill=green!5,opacity=.25] (C.center) to (B.center) to (B2.center) to (C2.center) to (C.center);
\path[fill=green!5,opacity=.25] (O2.center) to (A2.center) to (B2.center) to (C2.center) to (O2.center);
\path[fill=gray!10,opacity=.25] (O1.center) to (A1.center) to (B1.center) to (C1.center) to (O1.center);

%\node[draw=none,shape=circle,fill, inner sep=2pt] (d1) at (0,1,0){};  % circle

\node (P1) at (1.5,3,\zlevel) {};
\node (P2) at (2.5,3,2.75)  {};
\node (P3) at (1.5,3,2.75)  {};
\draw[-latex,orange] (P1.center) -- node [right] {$V$}	 (P2.center);
\draw[-latex,orange!85] (P1.center) -- node [left] {$\sigma(V)$} (P3.center);
\draw[-latex,dashed,line width=0.5pt,dashed] (P2.center) -- node [above] {$\sigma$} (P3.center);

\node[label={270:{$0(e)$}},circle,fill,inner sep=1pt] (0e) at (1.5,3,\zlevel) {};
\node[label={270:{$e$}},circle,fill,inner sep=1pt] (e) at (1.5,3,0)  {};
\node[label={270:{$x$}},circle,fill,inner sep=1pt] (x) at (0,3,0)  {};

\path[fill=green!5,opacity=.25] (B.center) to (A.center) to (A2.center) to (A2.center) to (B.center);
\path[fill=green!5,opacity=.25] (A.center) to (O.center) to (O2.center) to (O2.center) to (A.center);

    \end{tikzpicture}
    \end{document}

    	%Curva di Partenza
    \draw[red,line width=1pt] (1 ,0,\zlevel)
    .. controls (1.5,0.75,\zlevel)  and (2.5,0.5,\zlevel) .. (1.5,1.5,\zlevel)
		\foreach \p in {0,2,...,10}{						%punti per ottenere il vettore variazione
      		node[sloped,inner sep=0cm,above,pos=\p/10,
      			anchor=south west]
      			(Na \p){}
  		}
  	.. controls  (0.5,2.5,\zlevel) .. (P1.center)
		\foreach \p in {0,2,...,10}{						%punti per ottenere il vettore variazione
      		node[sloped,inner sep=0cm,above,pos=\p/10,
      			anchor=south west]
      			(Ma \p){}
  		}
  		;

  	%Curva Variata
    \draw[gray,line width=0.5pt,dotted] (1 ,0,\zlevel)
    .. controls (1.95,0.75,1.75)  and (2.8,0.5,2.5) .. (2.25,1.5,2.15)
		\foreach \p in {0,2,...,10}{						%punti per ottenere il vettore variazione
      		node[sloped,inner sep=0cm,above,pos=\p/10,
      			anchor=south west]
      			(Nb \p){}
  		}
  	.. controls  (1.5,2.5,\zlevel) .. (P2)
		\foreach \p in {0,2,...,10}{						%punti per ottenere il vettore variazione
      		node[sloped,inner sep=0cm,above,pos=\p/10,
      			anchor=south west]
      			(Mb \p){}
  		}
  		;
   	%Lo Jacobi Field
    \foreach \p in {0,2,...,8} {
      \draw[-latex,gray,line width=0.65pt,dotted] (Na \p.center) -- (Nb \p.center);
      \draw[-latex,gray,line width=0.65pt,dotted] (Ma \p.center) -- (Mb \p.center);    }